Revisiting~the~Classics: Online~RL~in~the~Programmable~Dataplane

### 2021-06-02 paper/paper.tex
Related Work.,..

@@

Two main firmware models govern how the computation-intensive parts of these tasks (tile-coding, action preference list construction, policy updates) are carried out:
\begin{description}
	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each performs its work sequentially. Computing an action list requires a \emph{read lock} on the policy. If an update occurs, the core requests a \emph{write lock} before updating, greatly limiting online throughput. \emph{Tile lists} are stored for update computation.
	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate on processing state vectors, minimising latency. Minions have a fixed list of work items, while a controller thread sends compute/update commands before awaiting worker completion. Work items are disjoint, requiring no policy locks. \emph{State vectors} are stored for update computation.
\end{description}
Each offers a different point of optimisation; if updates are disabled, then the \emph{\indfw{}} model can maximise throughput, while the \emph{\coopfw{}} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
We note that our parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is made possible only by the use of quantisation to integers, as atomic operations on floating point numbers typically do not exist outside of graphics- or tensor-processing hardware.

@@

\subsection{Intra-Agent Communication}
Even with embarrassingly parallel problems, optimising for latency requires meticulous care in how work is passed out and aggregated.
This is truer still when moving from the moderately fine-grained, efficient control of classical RL methods ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
Ordinarily, the serialisation/marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
On-chip execution and the nature of action preference computation allow us to use lockless atomic aggregation, removing the overheads of explicit messaging/packetisation.
Moreover, physically adjacent functional units may have special-purpose shared registers or share a small fast cache to accelerate communication further.

@@

\subsection{Work Allocation}
Due to the lack of dynamic memory allocation and to simplify action-value lookups, policies cannot be effectively stored sparsely.
As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in memory regions large enough to allocate space for them.
As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---ranging over CLS, CTM and IMEM in increasing latency.

We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded context of the least loaded ME.
The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weight the total cost per ME to account for the number of minion threads available.
This weighting specifically accounts for the presence of the controller thread on the first ME.
This work allocation and any cached per-work-item values are recomputed when any configuration is installed or changed.
Naturally, for $n$ tilings/tasks and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per task, storing $m/8$ and $\le8$ costs respectively.

@@

\section{Related Work}

?? NOTES: try to strengthen emphasis on ``other works can't do online learning in the dataplane''. This is absolutely, \SI{100}{\percent} the killer feature.

?? A switch/NIC architecture like \textcite{DBLP:conf/hotnets/StephensAS18} would remove the need for hard-coded island-island relationships.

?? Event-driven model could better support many timer-driven RL use-cases \textcite{DBLP:conf/hotnets/IbanezABM19}.

?? Taurus. They can't train online though!~\parencite{DBLP:journals/corr/abs-2002-08987}

?? Do switches~\parencite{DBLP:conf/hotnets/XiongZ19}.
?? Train offline, classical methods targetting all match-action P4 systems (no externs etc).

\parencite{DBLP:journals/corr/abs-2002-08987,DBLP:journals/corr/abs-2009-02353,DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731,langlet-ml-netronome}

?? N3IC run NNs on NIC~\parencite{DBLP:journals/corr/abs-2009-02353}, allow packet tagging to occur based on pre-trained BNN~\parencite{DBLP:conf/nips/HubaraCSEB16}.
?? both NFP (\SI{45}{\micro\second}) and NetFPGA (\SI{0.3}{\micro\second}).
?? We handle larger inputs (versus their \SI{256}{\bit}), offer online training.
?? Use our 5-dim as a point of comparison?

?? Or as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}.
?? Binary NNs ``more suitable for network hardware''~\parencite{DBLP:journals/corr/MiyashitaLM16}?

?? Can we collect some papers on per-packet ML/RL? Per-packet classification using NNs to build decision trees~\parencite{DBLP:conf/sigcomm/LiangZJS19}

?? RL for congestion control. ~\parencite{DBLP:journals/corr/abs-1910-04054}

?? Arbitrary training enhanced by in-network? General NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18}, RL-specfic~\parencite{DBLP:conf/isca/LiLYCSH19}.

?? Mechanisms to support full in-switch DDoS detection~\cite{tnms-ddos-victim-ident}.

?? Simialr HW has been used to get NN \emph{inference} \textcite{langlet-ml-netronome}---although \emph{actual NN latency} in his Fig 4.4 (IPG is inter-packet gap: his is not truly async but achieves perf by lying on the packet path. Note that 0 IPG murders his latencies).

@@

?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors!

>>>>

We present \emph{\approachshort{}}---\approach{}---to the best of our knowledge the first online RL in the dataplane...

@@

Two main firmware models govern how the computation-intensive parts of these tasks (tile-coding, action preference list construction, policy updates) are carried out:
\begin{description}
	\item[\Indfw{} (\cref{fig:single-and-parallel:single})] Separate threads listen for new states, and each performs its work sequentially. Computing an action list requires a \emph{read lock} on the policy. If an update occurs, the core requests a \emph{write lock} before updating, greatly limiting online throughput. \emph{Tile lists} are stored for update computation.
	\item[\Coopfw{} (\cref{fig:single-and-parallel:parallel,alg:parsa})] Threads cooperate on processing state vectors, minimising latency. Minions have a fixed list of work items, while a controller thread sends compute/update commands before awaiting worker completion. Work items are disjoint, requiring no policy locks. \emph{State vectors} are stored for update computation.
\end{description}
Each offers a different point of optimisation; if updates are disabled, then the \emph{\indfw{}} model can maximise throughput, while the \emph{\coopfw{}} model is designed to minimise decision latency and needs no locks for updating the policy (increasing \emph{online learning} throughput).
We note that our parallel, wait-free Sarsa implementation---\emph{ParSa} (\cref{alg:parsa})---is made possible only by the use of quantisation to integers, as atomic operations on floating point numbers typically do not exist outside of graphics- or tensor-processing hardware.
ParSa is described here at a high level, as an algorithm suited for \emph{any multiprocessor environment}.
We detail our communication primitives in \cref{sec:intra-agent-communication}, and our work scheduling strategy in \cref{sec:work-allocation}, eliding the details of tile-coding as they are well-understood.

@@

\subsection{Intra-Agent Communication}\label{sec:intra-agent-communication}
Even with embarrassingly parallel problems, optimising for latency requires meticulous care in how work is passed out and aggregated.
This is truer still when moving from the moderately fine-grained, efficient control of classical RL methods ($\sim$\SI{1}{\milli\second}) to its logical limit (tens of \si{\micro\second}).
Ordinarily, the serialisation/marshalling of requests, responses, and safe shared data access can incur significant extra overheads.
On-chip execution and the nature of action preference computation allow us to use lockless atomic aggregation, removing the overheads of explicit messaging/packetisation.
Moreover, physically adjacent functional units may have special-purpose shared registers or share a small fast cache to accelerate communication further.

@@

\subsection{Work Allocation}\label{sec:work-allocation}
Due to the lack of dynamic memory allocation and to simplify action-value lookups, policies cannot be effectively stored sparsely.
As a consequence, tiling space requirements scale exponentially with dimension count, and so higher-dimension tilings must be placed in memory regions large enough to allocate space for them.
As a result, tilings are split across memory regions, giving different access and compute costs to different tasks---ranging over CLS, CTM and IMEM in increasing latency.

We use a simple first-fit work placement algorithm run on the NFP, placing the largest work item into the least loaded context of the least loaded ME.
Each work item is a separate \emph{tiling} over a number of dimensions, noting that many such tilings may be grouped into a set (using different offsets for smoother coverage of the state space).
The cost of any individual work item (according to its dimension count and memory location) was empirically measured offline, and we weight the total cost per ME to account for the number of minion threads available.
This weighting specifically accounts for the presence of the controller thread on the first ME.
This work allocation and any cached per-work-item values are recomputed when any configuration is installed or changed.
Naturally, for $n$ tilings/tasks and $m$ threads this procedure is $\mathcal{O}{\left(n\log{m}\right)}$: two find/update min operations into binary heaps per task, storing $m/8$ and $\le8$ costs respectively.

@@

\section{Related Work}
\fakepara{In-network ML}
Taurus~\parencite{DBLP:journals/corr/abs-2002-08987} proposes that efficient line-rate, low-latency per-packet inference can be achieved by placing a run-time configurable grid topology of map-reduce processing units into the packet processing pipeline (implementing CNNs, LSTMs, and some classical approaches such as SVMs).
Their CGRA-based implementation is able to be repurposed at runtime and achieves sub-\si{\micro\second} extra latency, though online model training is not considered.

\emph{IIsy}~\parencite{DBLP:conf/hotnets/XiongZ19} shows how \emph{classical ML inference} techniques (SVMs, Na\"{i}ve Bayes, Decision Trees, K-Means) can be converted into match-action table structures which are compatible with \emph{any} P4 deployment.
They achieve mean \SI{2.62}{\micro\second} extra in-path latency on NetFPGA devices (at line rate in most cases), but can only investigate the use of pre-trained models.

A recent line of research in the community has been to investigate the use of \emph{Binarised/Bitwise Neural Networks} (BNNs)~\parencite{DBLP:conf/nips/HubaraCSEB16,DBLP:journals/corr/KimS16,DBLP:journals/corr/MiyashitaLM16} for line-rate, low latency packet classification.
\emph{BaNaNa SPLIT} shows the initial steps of this trend investigating their use as an offload mechanism~\parencite{DBLP:conf/sigcomm/SanvitoSB18,DBLP:journals/corr/abs-1801-05731}; DNN inference is often carried out on the \emph{CPU} in an effort to reduce latency imposed by batching and bulk transfer to the GPU.
However, fully-connected layers are cache-unfriendly on commodity CPUs~\parencite{DBLP:conf/sigcomm/SanvitoSB18}, with the hypothesis put forward that networking hardware would be a better fit for some parts of the inference pipeline.
This line of work has moved onto full, in-network packet tagging and classification by pre-trained BNNs with \emph{N3IC}~\parencite{DBLP:journals/corr/abs-2009-02353}.
N3IC achieves packet inference in \SI{45}{\micro\second} on the NFP, and \SI{0.3}{\micro\second} on NetFPGA for \SI{256}{\bit} inputs.
Comparatively, \approachshort{}-\Coopfw{} can process an identically-sized input (8-dim) in a median \SI{13.83}{\micro\second}.
Our work handles larger inputs (\SI{640}{\bit}) at lower latencies of \SI{34}{\micro\second}, and offers online learning.
We expect that a NetFPGA implementation of \approachshort{} would enjoy a similar factor of speedup.
No works investigate the runtime training of BNN function approximators in-NIC.

\textcite{langlet-ml-netronome} has shown the viability of NN inference using \SI{64}{\bit} quantised weights and inputs on NFP hardware, using \emph{in-path} compute rather than our asynchronous model.
Inference latency on small networks (3 layers, \num{30} layers) can be as high as \SI{500}{\micro\second} on line rate traffic, stressing the value of the asynchronous compute and quantisation we adopt.

We note that none of the other approaches listed here (or that we have seen) tackle the issue of \emph{online learning and control} in-network---we believe \approachshort{} has broken new ground in this regard.

\fakepara{In-network ML \emph{acceleration}}
Optimisation of distributed neural network training is an area where in-network compute has been key in the case of general NNs~\parencite{DBLP:conf/micro/LiPAYQPWSEK18} and RL-specific procedures~\parencite{DBLP:conf/isca/LiLYCSH19} using NetFPGAs to introduce floating point adders to process dedicated packet classes.
Distributed training is a partition-aggregate workload, where gradient vectors are sent back to a controller to update the `true' model---in-NIC processing allows these to be aggregated \emph{in-network}, overcoming incast behaviour and host bottlenecks.

\fakepara{RL for network control}
\Textcite{DBLP:conf/sigcomm/LiangZJS19} use deep RL techniques on a host machine to train an agent which can build efficient decision tree packet classifiers.
Their approach, \emph{NeuroCuts}, is able to produce time/space efficient classifiers for use in constrained environments (e.g., network hardware) beyond existing heuristic solutions.

Deep RL techniques have been used for QUIC congestion control optimisation~\parencite{DBLP:journals/corr/abs-1910-04054}.
A key facet of this work is the need for asynchronous RL in networks, where pauses for DNN-based inference can significantly harm throughput.

\fakepara{PDP design for asynchronous compute}
\Textcite{DBLP:conf/hotnets/StephensAS18} have proposed \emph{PANIC}, which places a routing fabric between distinct packet/data processing elements \emph{in a SmartNIC}.
Such designs would more generally enable asynchronous and novel compute in SmartNICs and programmable switches, for instance offering consistent and easy to use communication between workers versus hard-coded ME relationships.

Event-driven versions of the P4 pipeline have been suggested~\parencite{DBLP:conf/hotnets/IbanezABM19}, 
Timer events and device state changes would empower in-network RL use-cases, signalling timesteps for RL agents or a new, effective, fine-grained sources of input state.

@@

?? Future directions for community: want to see more general-purpose async compute on-chip! Larger form-factors! More asynchrony!
